---
title: "ST_558_NHL_Project"
author: "Pramodini Karwande"
date: "6/13/2021"
output: 
    html_document: default  
    github_document: default
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = FALSE , eval = TRUE, warning = FALSE, message = FALSE)
```
```{r }
# section that notes the required packages needed to run the code
library("httr")
library("jsonlite")
library("dplyr")
library("kableExtra")
library("knitr")
```


# NHL records API
###You should write functions to contact the NHL records API for the endpoints listed below. The functions should return well-formatted, parsed data (usually a data frame). Where possible, the user should have the option to specify the franchise of choice by both name and ID number - you’ll need to map the names to ID numbers yourself.
```{r }
#Grab base URL for API
base_url <- "https://records.nhl.com/site/api"
urlExtention <- "?cayenneExp=franchiseId="
reqID <- ""
team_stat_url <- "https://statsapi.web.nhl.com/api/v1/teams"
team_extend_url <- "?expand=team.stats"

getNhlData<- function(franchiseID=NULL, franchiseName=NULL, tab_name=NULL) {
reqID <-  getID(franchiseID, franchiseName, tab_name)

if(tab_name == "franchise-detail")
{
  urlExtention <- "?cayenneExp=mostRecentTeamId="
}

if(!is.null(reqID)) {
  full_url <- paste0(base_url, "/", tab_name,urlExtention,reqID)
}
else {
  if((tab_name == "franchise-team-totals") || (tab_name == "franchise") ) {
    full_url <- paste0(base_url, "/", tab_name)
  }
}

# GET request to API
res <- GET(full_url)

franchise_data <- fromJSON(rawToChar(res$content), flatten = TRUE)
franchise_data <- data.frame(franchise_data)

if(tab_name == "franchise") {
  franchiseDataTab <- franchise_data %>% mutate("Team"=  paste0(data.teamPlaceName," ",data.teamCommonName)) %>% select("ID" = data.id, "First Season Id"= data.firstSeasonId, "Last Season Id"= data.lastSeasonId, Team)
  #franchiseDataTab <- na.omit(franchiseDataTab)
  knitr::kable(franchiseDataTab, caption = "NHL Data") %>% kable_styling()
}
else {
  knitr::kable(franchise_data, caption = "NHL Data") %>% kable_styling()
}
}


getID <- function(franchiseID=NULL, franchiseName=NULL, tab_name=NULL){
  retVal <- NULL
  if (!is.null(franchiseID)) {
    retVal<- franchiseID
    return(retVal)
  }
  else if (!is.null(franchiseName)) {
    retVal <- getName(franchiseName)
    return(retVal)
  }
  else {
    return(retVal)
  }
}

getName <- function(franchiseName=NULL) {
    tab_name <- "franchise"
    full_url <- paste0(base_url, "/", tab_name)
    res <- GET(full_url)

    franchiseNameData <- fromJSON(rawToChar(res$content), flatten = TRUE)
    franchiseNameData <- data.frame(franchiseNameData)
    franchiseID <- franchiseNameData %>% filter(data.fullName == franchiseName) %>% select(data.id)

    return(franchiseID)
} 


```


# NHL stats API
###You should write a function to contact the NHL stats API for the ?expand=team.stats modifier. The function should be able to take a single team or return data from all teams.
```{r}
getNHLTeamStatsData <- function(teamID=NULL) {
  if (!is.null(teamID)) {
    team_stat_full_url <- paste0(team_stat_url,"/",teamID,team_extend_url)
  }
  else {
    team_stat_full_url <- paste0(team_stat_url,team_extend_url)
  }
  
  res <- GET(team_stat_full_url)

    team_stat_data <- fromJSON(rawToChar(res$content), flatten = TRUE)
    team_stat_data <- data.frame(team_stat_data)
    knitr::kable(team_stat_data, caption = "NHL Team Stat Data") %>% kable_styling()
}

# *****************TESTING START***********************
 getNHLTeamStatsData(NULL)
 getNHLTeamStatsData(3)
# *****************TESTING END***********************
```
#Wrapper Function
###You should write a wrapper function that is essentially a one-stop-shop for the user to access any of the API endpoints you did above. That is, this function should simply call the appropriate endpoint as per the users request (including any modifiers, teamIDs, etc.)
```{r }

nhlWrapper <- function(nhlDataRequest, franchiseID=NULL,franchiseName=NULL, tab_name=NULL) {
  output = switch (nhlDataRequest,
    "/franchise"                = getNhlData(franchiseID,franchiseName,"franchise"),
    "/franchise-team-totals"    = getNhlData(franchiseID,franchiseName,"franchise-team-totals"),
    "/franchise-season-records" = getNhlData(franchiseID,NULL,"franchise-season-records"),
    "/franchise-goalie-records" = getNhlData(franchiseID,NULL,"franchise-goalie-records"),
    "/franchise-skater-records" = getNhlData(franchiseID,NULL,"franchise-skater-records"),
    "/franchise-detail"         = getNhlData(franchiseID,NULL,"franchise-detail"),
    "/individual-team-stat"     = getNHLTeamStatsData(teamID = 2),
    "/teams-stat"               = getNHLTeamStatsData(NULL),
    )
  return(output)
}

# *****************TESTING START***********************

nhlWrapper("/franchise",NULL, NULL, "franchise" )
nhlWrapper("/franchise-team-totals",NULL, NULL, "franchise-team-totals" )
nhlWrapper("/franchise-season-records",3, NULL, "franchise-season-records" )
nhlWrapper("/franchise-goalie-records",8, NULL, "franchise-goalie-records" )
nhlWrapper("/franchise-skater-records",4, NULL, "franchise-skater-records" )
nhlWrapper("/franchise-detail",2, NULL, "franchise-detail" )
#getNhlData(NULL, "Montreal Wanderers", "franchise")
#getNhlData(6, "Montreal Wanderers", NULL)
#getNhlData(NULL, NULL, NULL)

# *****************TESTING END***********************
```
# Exploratory Data Analysis (EDA)

### Data from at least two endpoints (possibly combining them into one)
```{r}
#Team Stat data
res <- GET(paste0(team_stat_url,team_extend_url))
teams_stat <- fromJSON(rawToChar(res$content), flatten = TRUE)
teams_stat <- data.frame(teams_stat) %>% rename("data.id"="teams.franchiseId")

# Franchise data
res <- GET(paste0(base_url, "/", "franchise"))
franchise_join_data <- fromJSON(rawToChar(res$content), flatten = TRUE)
franchise_join_data <- data.frame(franchise_join_data)


#Fanchise and Team stat joint data
franchise_team_stat <- inner_join(franchise_join_data, teams_stat, by = "data.id")
franchise_team_stat <- franchise_team_stat %>% select("franchiseID"="data.id","teams.id", "teams.firstYearOfPlay", "teams.active","teams.venue.city","teams.venue.id","teams.division.id","teams.division.name", "teams.conference.id", "teams.conference.name" ) %>% na.omit()
franchise_team_stat
```
## You should create at least two new variables that are functions of the variables from a data set you use
```{r}
res <- GET(paste0(paste0(base_url, "/", "franchise-goalie-records",urlExtention,12)))
goli_12_data <- fromJSON(rawToChar(res$content), flatten = TRUE) 
goli_12_data <- goli_12_data %>% data.frame(goli_12_data) %>% select(data.activePlayer, data.firstName, data.franchiseName, data.gamesPlayed, data.lastName, data.gameTypeId, data.playerId, data.mostGoalsAgainstOneGame, data.ties, data.wins, data.losses ) %>% mutate(winRate= round(data.wins/data.gamesPlayed, 2) , lossRate= round(data.losses/data.gamesPlayed,2))
goli_12_data

```
### You should create some contingency tables
```{r}

```






